---
title: "Exploration DataProducts ShinyProject"
author: "Sep"
date: "Friday, October 17, 2014"
output: html_document
---

# Exploration DataProducts ShinyProject
## For Developing Data Products - 17/10/2014
## Author "Sep"


# The general plan

## The Initial ideas

* Interactive fractal generation
* Some form of ODE systems - predator/prey maybe?
* Game theory project - prisoners dilema?
* percolalltion and forest fires <-- could be a real treat!
* Interactive statistics - plotting distributions
* Black Jack simulator?
* Interactive statistics - Using the General Social Survey Data 1972-2012 [Pick this]


## Why pick the given project?

* Good to get more experience working with data, will overlap with some of the 
other courses that I'm currently taking.
* I've already done most of the other projects in different languages.
* I want to know more about the given data!!

## Need Data?

Yes. The data I'm going to use is from the [ICPSR](http://www.icpsr.umich.edu/icpsrweb/ICPSR/studies/34802/version/1)
Here is [a link to the codebook](http://www.icpsr.umich.edu/cgi-bin/file?comp=none&study=34802&ds=1&file_id=1136502&path=ICPSR)

The data is on the large side, as such it will need cutting.

I'll be starting off with a cleaned data set - from another [Duke Uni] course I've been taking, I will however need to cut the 
data down even further so that it can be easily uploaded for use in the shiny app.

## Aim <span style="color:green">$\checkmark$</span>

Make a shiny app! Want to have a drop menus for 

* Plot type 
* variables of interest i.e Var 1 against Var 2 or Var 3 againsts Var 1 etc..
    * This will need a selection process for number of variables user is interested in
    if two vars you can have these plots.
    * sub section plotting? how about splitting plots by group depending on the users vars of interest?
    * something like: if box plot make subsectioning available?
* I want check box that when turned on outputs a summary of the chosen data cut by groupings.


## Prelim todos (and don'ts)

Philosophy for the project: Keep it small, high quality and interesting. Think 2 mins of use by an average user.

* explore the data and ID vars of interest.
    * keep the number of vars modest! This is a proof of concept not a full flashy product.
* produce a smaller data set for the purpose of this shiny app.
* Make a small map of what the UI should look like and what I want the user to see exactly
* Get the shiny app running locally
* Add documentation 
* Upload to github
* get a pitch presentation together about the app 5 slides.
    

# Some exploration of the data

Start by loading up the Duke cleaned gss data

```{r}
load("../data/statistics-project-gss.Rdata")
names(gss)
```


From the names, there are few that look interesting for analysis
* year
* age
* sex
* edu
* coninc
* religion

## subsetting the data

Make a data table with dplyr

```{r}
library(dplyr)
gsstbl <- tbl_df(gss[,c("year", "age", "sex", "educ","degree", "coninc", "relig")])
```

Have a look at the structure and summary of the data

```{r}
str(gsstbl)
summary(gsstbl)
```


## pairs plotting

Have a look at scatter plot pairing the data. As a note the pairs function can take some time
so instead take a random sample from our table

```{r}
num_rand_rows <- 1000 # randomly sample this many rows
gsstbl %>% sample_n(size=num_rand_rows) %>% pairs
```


## Look at some histograms of interest concerning income/gender

A barplot of the highest degree achieved

```{r}
## Alternative to dplyr piping
# barplot( table(gsstbl$degree) )

gsstbl %>% select(degree) %>% table %>% barplot
```

Bar plot of average family income by highest degree

```{r}
library(ggplot2)
gsstbl %>%
        group_by(degree) %>%
        summarize(meanFamilyIncome=mean(coninc,na.rm=TRUE)) %>%
        ggplot(aes(x=degree,y=meanFamilyIncome,fill=degree)) + 
        geom_bar(stat="identity") + 
        labs(title="Average income by highest form of education")
```


Q: how has the average family income changed over the years based on your degree

```{r}
gsstbl %>%
    group_by(year,degree) %>%
    summarize(meanInc=mean(coninc,na.rm=TRUE)) %>%
    ggplot(aes(x=year,y=meanInc, group=degree,color=degree)) + 
    geom_point() + 
    geom_line() + 
    labs(title="Change in time of average income grouped by degree")
```

Have a look at a facetted plot adding a linear model to the graphs

```{r}
gsstbl %>%
    group_by(year,degree,sex) %>%
    summarize(meanInc=mean(coninc,na.rm=TRUE)) %>%
    ggplot(aes(x=year,y=meanInc,color=degree)) + 
    geom_point() + 
    geom_line() + 
    facet_wrap(~degree, scales="free") +
    geom_smooth(method="lm") +
    labs(title="Change in time of average income grouped by degree with linear fit")
```

Have a look at the trend difference between men and women
```{r}
gsstbl %>%
    group_by(year,degree,sex) %>%
    summarize(meanInc=mean(coninc,na.rm=TRUE)) %>%
    ggplot(aes(x=year,y=meanInc,color=sex, group=sex, color=sex)) + 
    geom_point() + 
    geom_line() + 
    facet_wrap(~ degree, scales="free") +
    geom_smooth(method="lm", se=FALSE) +
    labs(title="Change in time of average income grouped by degree & gender with linear fit")
```


